Staged Abstract Interpreters

### 2018-11-04 paper/10_intro.tex
minor note

Futamura projections \cite{Futamura1999, futamura1971partial} reveal the close connection between 
compilers and interpreters. 
The first Futamura projection specifically shows that specializing an interpreter with respect to the input program yields 
an equivalent executable.
Multi-stage programming (MSP) \cite{taha1999multi, DBLP:conf/pepm/TahaS97} is a popular way to write 
generic programs that and generates specialized and efficient programs;  \note{TR: connection from Futamura to MSP doesn't flow well}
the generated program then can be executed at next stage. 
As an approach to partial evaluation, MSP requires explicit staging annotations from the implementers.
The staging annotations controls which part of the program should be specialized, and are considered 
as manual binding-time analysis. On the other hand, partial evaluation \cite{DBLP:books/daglib/0072559} 
can be considered as a two-stage (static or dynamic) instance of MSP but analyzing binding-time automatically. 
\note{TR: are last two sentences relevant? perhaps rephrase in the direction of ``here's who to implement
the futamura projection idea in practice''}

Recent research \note{TR: Cousot was already systematic. Both Futamura and Cousot are results from the 1970s -- build a connection?} uncovers a systematic approach to construct sound abstract interpreters 
from concrete interpreters \cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10, DBLP:journals/pacmpl/DaraisLNH17}.
Given the structural similarity between concrete and abstract interpreters, intellectually 
it is natural to raise the question whether it is possible to stage an abstract interpreter, and how 
to effectively specialize an abstract interpreter having considered their functional dissimilarity. 
This paper studies the application of the first Futamura projection on abstract interpreters, 
and the approach to construct efficient abstract interpreters by multi-stage programming.

@@

>>>>

Futamura projections \cite{Futamura1999, futamura1971partial} reveal the close connection between 
compilers and interpreters. 
The first Futamura projection specifically shows that specializing an interpreter with respect to the input program yields 
an equivalent executable.
Multi-stage programming (MSP) \cite{taha1999multi, DBLP:conf/pepm/TahaS97} is a popular way to write 
generic programs that and generates specialized and efficient programs;  \note{TR: connection from Futamura to MSP doesn't flow well}
the generated program then can be executed at next stage. 
As an approach to partial evaluation, MSP requires explicit staging annotations from the implementers.
The staging annotations controls which part of the program should be specialized, and are considered 
as manual binding-time analysis. On the other hand, partial evaluation \cite{DBLP:books/daglib/0072559} 
can be considered as a two-stage (static or dynamic) instance of MSP but analyzing binding-time automatically. 
\note{TR: are last two sentences relevant? perhaps rephrase in the direction of ``here's how to implement
the futamura projection idea in practice''}

Recent research \note{TR: Cousot was already systematic. Both Futamura and Cousot are results from the 1970s, but have
profited from more recent results (MSP and AAM)  -- build a connection?} uncovers a systematic approach to construct sound abstract interpreters 
from concrete interpreters \cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10, DBLP:journals/pacmpl/DaraisLNH17}.
Given the structural similarity between concrete and abstract interpreters, intellectually 
it is natural to raise the question whether it is possible to stage an abstract interpreter, and how 
to effectively specialize an abstract interpreter having considered their functional dissimilarity. 
This paper studies the application of the first Futamura projection on abstract interpreters, 
and the approach to construct efficient abstract interpreters by multi-stage programming.

@@

\note{TR: insert a commuting diagram here? concrete vs abstract,
general (interpreter) vs specialized (compiler). can combine both dimensions.}
